---
title: "DistributedLag_group_control-demand"
author: "Arpita"
date: '2023-05-11'
output: html_document
---

```{r}
#install.packages("lubridate")
library(data.table)
library(lubridate)
library(dplyr)
library(tidyr)
```

```{r}
#devtools::install_github("danielmork/dlmtree", ref="development")
library(coda)
library(ggplot2)
library(ggpubr)
library(dlmtree)
library(splines)
library(readxl)
library(foreach)
#vignette("nested")
```

```{r}
countries <- c('CAL', 'CAR', 'CENT', 'FlA', 'MIDA', 'MIDW', 'NE', 'NW', 'SE', 'SW', 'TEN', 'TEX')
if (!file.exists("tdlm_models")){
    dir.create("tdlm_models")
}
data_list <- list()
lag_data_list <- list()
lead_data_list <- list()
complete_list <- list()
for (i in (1:length(countries))){
  print(countries[i])
  data <- read_xlsx(paste("data/Region_", countries[i], ".xlsx", sep=""), guess_max = 70000)
  #print(nrow(data))
  setDT(data)
  
  # Use filters after the shifted solar data is created
  #data <- data[!is.na(`NG: SUN`)]
  #data <- data[!which(`NG: SUN`<0)]
  #data <- data[which(`CO2 Emissions Generated`>0)]
  
  data[, date2 := as_date(`Local date`)]
  data[, dow := wday(date2)]
  data[, week := week(date2)]
  data[, month := month(date2)]
  data[, year := year(date2)]
  data[, demand := `D`]
  data[, index := 1:.N]
  
  # Make column names safe for use in formulas
  setnames(data, make.names(names(data)))
  #print(colnames(data))
  
  # Only consider data between during time period
  start_date = "2018-07-01"
  end_date = "2023-07-01"
  data <- data[data$date2 >= start_date & data$date2 < end_date, ]
  
  #convert negative solar values and zero CO2 emissions as NA. 
  data$NG..SUN = replace(data$NG..SUN , which(data$NG..SUN  < 0), 0)
  data$NG..SUN = replace(data$NG..SUN , which(data$CO2.Emissions.Generated <= 0), NA)

  # Create lagged data
  lags <- 12
  lag_dat <- as.matrix(data[, shift(NG..SUN, 0:lags, type = "lag")])
  #lead_dat <- as.matrix(data[, shift(NG..SUN, 10:1, type = "lead")])
  #lag_dat <- cbind(lead_dat, lag_dat)
  # data[, demand_ma10 := rollmean(demand, k = 10)] -- CHECK! 
  # Complete data rows

  data_list[[i]]<-data #Note: this data may contain NA values
  lag_data_list[[i]]<- lag_dat #Note: this data may contain NA values
  #lead_data_list[[i]]<-lead_dat
  complete_list[[i]] <- which(complete.cases(lag_dat) & 
                            complete.cases(data[, .(CO2.Emissions.Generated, demand)]))

}
```


```{r}
# Individual Nonlinear lagged model

mlist <- lapply(countries, function(x) list())
gr <- lapply(countries, function(x) list())
names(mlist) <- countries
names(gr) <- countries
restarts <- 5


# Iterate through regions, assess convergence, increase # trees
for (i in 1:length(countries)) {
  for (ntrees in c(20, 40, 80)) {
    cat(countries[i], "ntrees =", ntrees, "\n")
    data <- data_list[[i]]
    complete <- complete_list[[i]]
    lag_dat <- lag_data_list[[i]]
    splits <- seq(quantile(lag_dat[complete,], .05), quantile(lag_dat[complete,], .95), length.out = 10) #5%and 95% percentile
    
    # Run monotone-TDLNM
    model_run <-
      foreach(m = 1:restarts,
              .errorhandling = "remove",
              .packages = c("dlmtree", "data.table", "splines"),
              .export = c("data", "complete", "splits", "lag_dat", "g0", "s0"),
              .verbose = F) %do%
      {
        set.seed(m)
        tdlnm(I(-1*CO2.Emissions.Generated) ~
                ns(index, df = 10) +
                ns(Hour, df = 5) +
                demand,
              data[complete],
              exposure.data = lag_dat[complete,],
              exposure.splits = splits,
              monotone = T, shrinkage = F, 
              monotone.sigma = diag(lags + 1) * 2.701^2,
              time.split.prob = lags:1,
              n.trees = ntrees,
              #n.burn=10000, n.iter=5000,
              verbose = F)
      }
    
    # Check convergence
    s <- lapply(model_run, summary, cenval = median(data_list[[i]][NG..SUN>0, NG..SUN]),
                mcmc = TRUE, verbose = FALSE)
    mcmc_mtdlnm <- do.call(mcmc.list, lapply(1:length(model_run), function(i) {
      d <- as.data.frame.table(s[[i]]$dlm_mcmc)
      setDT(d)
      mcmc(dcast(d, Var3 ~ Var1 + Var2, value.var = "Freq")[, Var3 := NULL][])
    }))
    gr_test <- gelman.diag(mcmc_mtdlnm, transform = F, autoburnin = F, multivariate = F)
    cat(" GR =", median(gr_test$psrf[, 1]), "\n")
    
    if (is.null(gr[[countries[i]]]$psrf) || (median(gr_test$psrf[, 1]) < median(gr[[countries[i]]]$psrf[, 1]))) {
      mlist[[countries[i]]] <- model_run
      gr[[countries[i]]] <- gr_test
    }
  }
  dm_mlist<- mlist[[countries[i]]]
  save(dm_mlist, file=paste("tdlm_models/mlist_",countries[i],".Rda",sep=""))
  dm<- combine.models(mlist[[countries[i]]])
  save(dm, file=paste("tdlm_models/dm_mod_",countries[i],".Rda",sep=""))
  ntrees_best <- gr[[countries[i]]]
  save(ntrees_best, file=paste("tdlm_models/gr_",countries[i],".Rda",sep=""))
}
```


```{r}
# Individual Nonlinear lagged model for neighboring effect analysis

#One country's CO2 reduction based on the "base" country
countries <- c('CAL', 'CAR', 'CENT', 'FlA', 'MIDA', 'MIDW', 'NE', 'NW', 'SE', 'SW', 'TEN', 'TEX')
base_countries =c(1,1,2,2,3,3,3,3,5,5,5,6,6,8,8,9,9,9,9,10,11,11,11,12)
other_countries =c(10,8,5,9,6,8,10,12,2,6,11,9,11,1,10,2,4,6,11,1,2,5,9,3)

combo_list<-list()
for (i in (1:length(base_countries))){
    base_country <- countries[base_countries[i]]
    other_country <- countries[other_countries[i]]
    combo_list<- append(combo_list, paste(base_country, other_country, sep="-"))
}
combo_list<- unlist(combo_list)

mlist <- lapply(combo_list, function(x) list())
gr <- lapply(combo_list, function(x) list())
names(mlist) <- combo_list
names(gr) <- combo_list
restarts <- 5
```

```{r}
run_analysis <- function(combo, base, other, i){
    print(paste("base=",base, ", other=", other, sep=""))
    data <- combo
    setDT(data)
    #print(colnames(data))
    #data <- data[!is.na(`NG: SUN`)]
    #print(nrow(data))
    #data <- data[!which(`NG: SUN`<0)]
    #print(nrow(data))
    #data <- data[which(`CO2 Emissions Generated`>=0)]
    #print(nrow(data))

    data[, date2 := as_date(`Local.date`)]
    data[, dow := wday(date2)]
    data[, week := week(date2)]
    data[, month := month(date2)]
    data[, year := year(date2)]
    data[, demand := `D`]
    data[, index := 1:.N]
    
    # Make column names safe for use in formulas
    setnames(data, make.names(names(data)))
    #print(colnames(data))
    
    
    # Create lagged data
    lags <- 12
    lag_dat <- as.matrix(data[, shift(NG..SUN, 0:lags, type = "lag")])
    #lead_dat <- as.matrix(data[, shift(NG..SUN, 10:1, type = "lead")])
    #lag_dat <- cbind(lead_dat, lag_dat)
    
    complete<- which(complete.cases(lag_dat) & 
                                complete.cases(data[, .(CO2.Emissions.Generated, demand, NG..SUN, NG..SUN.y)]))
    
    
    for (ntrees in c(20)) {
    cat(combo_list[i], "ntrees =", ntrees, "\n")
    splits <- seq(quantile(lag_dat[complete,], .05), quantile(lag_dat[complete,], .95), length.out = 10) #5%and 95% percentile
    
    # Run monotone-TDLNM
    model_run <-
      foreach(m = 1:restarts,
              .errorhandling = "remove",
              .packages = c("dlmtree", "data.table", "splines"),
              .export = c("data", "complete", "splits", "lag_dat", "g0", "s0"),
              .verbose = F) %do%
      {
        set.seed(m)
        tdlnm(I(-1*CO2.Emissions.Generated) ~
                ns(index, df = 10) +
                ns(Hour, df = 5) +
                demand,
              data[complete],
              exposure.data = lag_dat[complete,],
              exposure.splits = splits,
              monotone = T, shrinkage = F, 
              monotone.sigma = diag(lags + 1) * 2.701^2,
              time.split.prob = lags:1,
              n.trees = ntrees,
              #n.burn=10000, n.iter=5000,
              verbose = F)
      }
    
    # Check convergence
    s <- lapply(model_run, summary, cenval = median(data[NG..SUN>0, NG..SUN]),
                mcmc = TRUE, verbose = FALSE)
    mcmc_mtdlnm <- do.call(mcmc.list, lapply(1:length(model_run), function(i) {
      d <- as.data.frame.table(s[[i]]$dlm_mcmc)
      setDT(d)
      mcmc(dcast(d, Var3 ~ Var1 + Var2, value.var = "Freq")[, Var3 := NULL][])
    }))
    gr_test <- gelman.diag(mcmc_mtdlnm, transform = F, autoburnin = F, multivariate = F)
    cat(" GR =", median(gr_test$psrf[, 1]), "\n")
    
    if (is.null(gr[[combo_list[i]]]$psrf) || (median(gr_test$psrf[, 1]) < median(gr[[combo_list[i]]]$psrf[, 1]))) {
      mlist[[combo_list[i]]] <- model_run
      gr[[combo_list[i]]] <- gr_test
    }
  }
  dm_mlist<- mlist[[combo_list[i]]]
  save(dm_mlist, file=paste("tdlm_models/mlist_",combo_list[i],".Rda",sep=""))
  dm<- combine.models(mlist[[combo_list[i]]])
  save(dm, file=paste("tdlm_models/dm_mod_",combo_list[i],".Rda",sep=""))
  ntrees_best <- gr[[combo_list[i]]]
  save(ntrees_best, file=paste("tdlm_models/gr_",combo_list[i],".Rda",sep=""))
}
```

```{r}
for (i in (1:length(combo_list))){
  {
    {
    
      base_country <- countries[base_countries[i]]
      other_country <- countries[other_countries[i]]
      
      print(combo_list[i])
      
      #Load data of base_country
      d<- data_list[[base_countries[i]]]
      
      #Only keep the desired number of columns
      d <- d[, c("UTC.time", "NG..SUN")]
      
      #Create one merged file
      d<-merge(d, data_list[[other_countries[i]]], by=c("UTC.time"))
      print(colnames(d))
      #d$D<- d$D.y
      d$NG..SUN<- d$NG..SUN.x 
      #d$CO2.Emissions.Generated <- d$CO2.Emissions.Generated.y
      d <- d[, c("Local.date", "Hour", "D", "NG..SUN", "NG..SUN.y", "CO2.Emissions.Generated")]
      
      #Run analysis
      run_analysis(combo = d, base = base_country, other = other_country, i=i)
    }
  }
  
}


